Large Scale Structure Reconstruction with Short-Wavelength Modes

### 2020-02-08 paper/v2.tex
1

\end{abstract}
\maketitle
\noindent {\it Introduction.}
Measuring the distribution of matter on large scales is one of the goals of cosmological surveys. The information contained on large scales may provide information about issues ranging from the turnover in the power spectrum (and therefore the total matter density) to the accelerated universe to anomalies observed in the cosmic microwave background to primordial nongaussianity. Direct measurements are difficult because of observational systematic effects, so indirect approaches have been considered. Small scale structure grows differently in the presence of an large-scale overdensity: it is as if the mean background density is larger than on average. This relation between long- and short-wavelength modes has been discussed in recent years \cite{Baldauf:2011fer}\cite{Li:2014ssc}\cite{Barreira:2017res}. The method of using the small scale position-dependent power spectrum to compute the squeezed-limit bispectrum also indicates that small scale perturbations can be used to infer large scale information \cite{Chiang:2014pos}\cite{Chiang:2015poss}. In order to fully achieve this goal, here we construct a quadratic estimator to measure long-wavelength modes indirectly.

@@

\frac{\partial \delta(\vec{x},\tau)}{\partial \tau} &+&\vec{\nabla}\times [(1+\delta(\vec{x},\tau))\vec{v}(\vec{x},\tau)] =0\eql{fluid1} \\
(\frac{\partial}{\partial \tau} +\vec{v}(\vec{x},\tau)&\times&\vec{\nabla})\vec{v}(\vec{x},\tau)=-\frac{da}{d\tau}\frac{\vec{v}(\vec{x},\tau)}{a}-\vec{\nabla}\Phi \eql{eulereq}\\

@@

These equations fully determine the time evolution of the local density contrast $\delta$ and the peculiar velocity field $\vec{v}=d\vec{x}/d\tau$.

@@

The first order term ${\delta}^{(1)}$ corresponds to linear evolution. The linear power spectrum is given by this first order term via:

@@

Substituting the perturbative series \ec{pert1} and \ec{pert2} into the Fourier transformed fluid equations \ec{fluid1}-\ec{fluid3} leads to an expression for the second-order density contrast:

@@

F_{2}(\vk_1,\vk_2)=\frac{5}{7}+\frac{2}{7}\frac{(\vk_1\cdot \vk_2)^2}{k_1^2 k_2^2}+\frac{\vk_1\cdot \vk_2}{2k_1k_2}(\frac{k_1}{k_2}+\frac{k_2}{k_1}).\eql{f2}

@@

Note that \ec{pert1}, \ec{pert2} and \ec{f2} are completely accurate only in an Einstein-de Sitter universe. Nonetheless for related calculations in a $\Lambda$CDM universe, the difference is found to be negligible \cite{Takahashi:2008to}, and thus we use the expressions from \ec{pert1} and \ec{f2} throughout this work.

@@

When one of the wavenumbers in the 3-point function in \ec{integral} is very small, that mode can be considered as a background mode. The small scale modes evolve in the presence of whatever long wavelength modes happen to be present. Therefore, 

@@

\delta^{(1)}(\vk_{s}) \delta^{(1)}(\vk_{s}-\vk){\delta^{(1)}}(\vk) 

@@

&=&\langle {\delta^{(1)}}(\vk_{s}) {\delta^{(1)}}(\vk_{s}-\vk) \rangle {\delta^{(1)}}(\vk)  \vs

@@

\noindent{\it Quadratic Estimator.}\label{sec2}
We can now construct the quadratic estimator for long-wavelength modes starting from \ec{2pt} and summing over as many pairs as possible with weights that maximize the signal to noise. As with the case of CMB lensing, we can write the general form of the estimator by averaging over pairs of short-wavelength modes:
\begin{eqnarray}
\hat{\delta}^{(1)}(\vec{k}_l)=A(\vec{k}_l)\int \frac{d^3 \vec{k}_s}{(2\pi)^3} g(\vec{k}_s,\vec{k}_s'){\delta}(\vec{k}_s){\delta}(\vec{k}_s') \eql{quadest},
\end{eqnarray} 
with $g$ being a weighting function, $\vk_s'=\vk_l-\vk_s$ and $A$  defined by requiring that $\langle \hat{\delta}^{(1)}(\vec{k}_l) \rangle={\delta}^{(1)}(\vec{k}_l)$:
\begin{eqnarray}
A(\vec{k}_l)=\bigg[\int \frac{d^3 \vec{k}_s}{(2\pi)^3} g(\vec{k}_s,\vec{k}_s')f(\vec{k}_s,\vec{k}_s')  \bigg]^{-1} \eql{a}
\end{eqnarray}
In the absence of shot noise the Gaussian noise is given by:
\be 
\langle \hat{\delta}^{(1)}(\vk_{l})\hat{\delta}^{(1)*}(\vk_{l}') \rangle = (2\pi)^3 \delta_{\rm D}(\vk_{l}-\vk_{l}')[P_{\rm lin}(k_{l})+N(\vk_l)]
\ee 
with 
\begin{eqnarray}
&&N(\vec{k}_{l})=2A^2(\vk_{l})\vs
&&\times\int \frac{d^3 \vec{k}_{s}}{(2\pi)^3} g^2(\vec{k}_{s},\vk_l-\vec{k}_{s})P_{\rm nl}(k_{s})P_{\rm nl}(|\vk_l-\vk_s|)
\end{eqnarray}
where $P_{\rm nl}$ is the nonlinear power spectrum. Minimizing the noise term we can fix the form of $g$ to be:
\begin{eqnarray}
g(\vec{k}_{s},\vec{k}_{s}')=\frac{f(\vec{k}_{s},\vec{k}_{s}')}{2P_{\rm nl}(k_{s})P_{\rm nl}(k_{s}')}
\end{eqnarray} 
The noise term reduces simply to $N(\vk_l)=A(\vk_l)$. We find by testing that the value of $N$ is very insensitive to the choice of the lower limit of the integration \ec{a}, since most of the contribution comes from large $k_s$.

>>>>

\end{abstract}
\maketitle
\noindent {\it Introduction.}
Measuring the distribution of matter on large scales is one of the goals of cosmological surveys \cite{LSST:2012ls}\cite{Wfirst:2012jg}. The information contained on large scales may provide information about issues ranging from the turnover in the power spectrum (and therefore the total matter density) to the accelerated universe to anomalies observed in the cosmic microwave background to primordial nongaussianity. Direct measurements are difficult because of observational systematic effects, so indirect approaches have been considered. Small scale structure grows differently in the presence of an large-scale overdensity: it is as if the mean background density is larger than on average. This relation between long- and short-wavelength modes has been discussed in recent years \cite{Masui:2010prim}\cite{Baldauf:2011fer}\cite{Jeong:2012foss}\cite{Li:2014ssc}\cite{Zhu:2016tidal}\cite{Barreira:2017res}\cite{Zhu:2018cm}. The method of using the small scale position-dependent power spectrum to compute the squeezed-limit bispectrum also indicates that small scale perturbations can be used to infer large scale information \cite{Chiang:2014pos}\cite{Chiang:2015poss}. In order to fully achieve this goal, here we construct a quadratic estimator to measure long-wavelength modes indirectly.

@@

\frac{\partial \delta(\vec{x},\tau)}{\partial \tau} &+&\vec{\nabla}\cdot [(1+\delta(\vec{x},\tau))\vec{v}(\vec{x},\tau)] =0\eql{fluid1} \\
\bigg[\frac{\partial}{\partial \tau} +\vec{v}(\vec{x},\tau)\cdot&\vec{\nabla}& \bigg]\vec{v}(\vec{x},\tau)=-\frac{da}{d\tau}\frac{\vec{v}(\vec{x},\tau)}{a}-\vec{\nabla}\Phi \eql{eulereq}\\

@@

\peikai{Here $a$ is the cosmological scale factor, $\Phi$ is the 3D gravitational potential and $\bar{\rho}_{\rm m}(a)$ is the mean matter density.} These equations fully determine the time evolution of the local density contrast $\delta$ and the peculiar velocity field $\vec{v}=d\vec{x}/d\tau$.

@@

The first order term ${\delta}^{(1)}$ corresponds to linear evolution. The linear power spectrum is given by this first order term via \peikai{averaging over modes in Fourier space}:

@@

\peikai{Here $\delta_{\rm D}$ is the Dirac delta function.} Substituting the perturbative series \ec{pert1} and \ec{pert2} into the Fourier transformed fluid equations \ec{fluid1}-\ec{fluid3} leads to an expression for the second-order density contrast:

@@

F_{2}(\vk_1,\vk_2)=\frac{5}{7}+\frac{2}{7}\frac{(\vk_1\cdot \vk_2)^2}{k_1^2 k_2^2}+\frac{\vk_1\cdot \vk_2}{2k_1k_2}\bigg[\frac{k_1}{k_2}+\frac{k_2}{k_1}\bigg].\eql{f2}

@@

Note that \ec{pert1}, \ec{pert2} and \ec{f2} are completely accurate only in an Einstein-de Sitter universe \peikai{and also assuming the case of a pressureless perfect fluid}. Nonetheless for related calculations in a $\Lambda$CDM universe, the difference is found to be negligible \cite{Takahashi:2008to}, and thus we use the expressions from \ec{pert1} and \ec{f2} throughout this work.

@@

When one of the wavenumbers in the 3-point function in \ec{integral} is very small, that mode can be considered as a background mode. The small scale modes evolve in the presence of whatever long wavelength modes happen to be present. Therefore,  \peikai{we can take the long-wavelength mode out of the bracket:}

@@

\delta^{(1)}(\vk_{s}) \delta^{(1)}(\vk_{s}'-\vk){\delta^{(1)}}(\vk) 

@@

&=&\langle {\delta^{(1)}}(\vk_{s}) {\delta^{(1)}}(\vk_{s}'-\vk) \rangle {\delta^{(1)}}(\vk)  \vs

@@

\noindent{\it Quadratic Estimator.}\label{sec2}
We can now construct the quadratic estimator for long-wavelength modes starting from \ec{2pt} and summing over as many pairs as possible with weights that maximize the signal to noise. As with the case of CMB lensing, we can write the general form of the estimator by averaging over pairs of short-wavelength modes:
\begin{eqnarray}
\hat{\delta}^{(1)}(\vec{k}_l)=A(\vec{k}_l)\int \frac{d^3 \vec{k}_s}{(2\pi)^3} g(\vec{k}_s,\vec{k}_s'){\delta}(\vec{k}_s){\delta}(\vec{k}_s') \eql{quadest},
\end{eqnarray} 
with $g$ being a weighting function, $\vk_s'=\vk_l-\vk_s$ and $A$ \peikai{is the normalization factor} defined by requiring that $\langle \hat{\delta}^{(1)}(\vec{k}_l) \rangle={\delta}^{(1)}(\vec{k}_l)$:
\begin{eqnarray}
A(\vec{k}_l)=\bigg[\int \frac{d^3 \vec{k}_s}{(2\pi)^3} g(\vec{k}_s,\vec{k}_s')f(\vec{k}_s,\vec{k}_s')  \bigg]^{-1} \eql{a}
\end{eqnarray}
In the absence of shot noise the Gaussian noise is given by:
\be 
\langle \hat{\delta}^{(1)}(\vk_{l})\hat{\delta}^{(1)*}(\vk_{l}') \rangle = (2\pi)^3 \delta_{\rm D}(\vk_{l}-\vk_{l}')[P_{\rm lin}(k_{l})+N(\vk_l)]
\ee 
with 
\begin{eqnarray}
&&N(\vec{k}_{l})=2A^2(\vk_{l})\vs
&&\times\int \frac{d^3 \vec{k}_{s}}{(2\pi)^3} g^2(\vec{k}_{s},\vk_l-\vec{k}_{s})P_{\rm nl}(k_{s})P_{\rm nl}(|\vk_l-\vk_s|)
\end{eqnarray}
where $P_{\rm nl}$ is the nonlinear power spectrum. Minimizing the noise term we can fix the form of $g$ to be:
\begin{eqnarray}
g(\vec{k}_{s},\vec{k}_{s}')
&=&\frac{f(\vec{k}_{s},\vec{k}_{s}')}{2P_{\rm nl}(k_{s})P_{\rm nl}(k_{s}')}\vs
&=&\frac{F_2(-\vec{k}_s,\vec{k}_s+\vec{k}_s')P_{\rm lin}(k_s)+F_2(-\vec{k}_s',\vec{k}_s+\vec{k}_s')P_{\rm lin}(k_s')}{P_{\rm nl}(k_{s})P_{\rm nl}(k_{s}')}\vs 
\end{eqnarray} 
The noise term reduces simply to $N(\vk_l)=A(\vk_l)$. We find by testing that the value of $N$ is very insensitive to the choice of the lower limit of the integration \ec{a}, since most of the contribution comes from large $k_s$.
